They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-15 paper/paper.tex
i -> r, new table

The general formula for the probability \(P_E(i)\) that a best-of-\(s\)
series lasts \(i\) rounds given constant probability \(p\) of one
team\footnote{If \(p > 0.5\), then I might say that this team is the
  better team (known omnisciently).} winning each round is

@@

\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(i) = \frac{(i - 1)!}{s_1!(i - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq:series-length}\end{equation}

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected
probabilities for \(s = 11\) under the naive assumption that
\(p_0 = 0.5\) for a given round, along with the observed proportions
\(P_O(i)\) in CoD SnD.

\begin{longtable}{rrrr}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds ($P_E(i)$, where $i \in [6, 7, 8, 9, 10, 11]$) under the assumption that each team has a 50\% probability ($p_0 = 0.5$) of winning each game. The observed frequencies observed are shown as a count $N_O(i)$ and as a proportion $P_O(i)$ of all series.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
\toprule
&  & \multicolumn{2}{c}{Observed} \\ 
\cmidrule(lr){3-4}
Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
\midrule
6 & $3.1\%$ & $4.7\%$ & 40 \\ 
7 & $9.4\%$ & $11.9\%$ & 101 \\ 
8 & $16.4\%$ & $16.5\%$ & 141 \\ 
9 & $21.9\%$ & $21.7\%$ & 185 \\ 
10 & $24.6\%$ & $21.5\%$ & 183 \\ 
11 & $24.6\%$ & $23.7\%$ & 202 \\ 
\bottomrule
\end{longtable}

@@

\begin{equation}\protect\hypertarget{eq:series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(i) - P_E(i))^2}{P_E(i)}, i \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}

@@

\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & \(p\) & \(\chi^2\) (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & \(p\) & \(\chi^2\) (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}

Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new
\(P_E(i)\) when re-applying Equation \ref{eq:series-length} for the
maximum likelihood estimate \(p_2 = 0.575\).\footnote{The method of
  moments and minimum \(\chi^2\) estimates for \(p\) are omitted simply
  because the results would be nearly identical to those for the maximum
  likelihood estimate of \(p\) (since they are all \(\approx 0.575\).} I
observe that \(P_E(i)\) is notably larger for \(p_{2}\) when
\(i \in [6, 7]\), more closely matching \(P_O(i)\). \(p_{2}\) is also
closer to \(P_O(i)\) for \(i \in [9, 10]\), although not for
\(i \in [8, 11]\).

\begin{longtable}[]{@{}rrrr@{}}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds under the naive assumption $p_0 = 0.5$ and the maximum likelihood estimation $p_2 = 0.575$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
Series lasts \(i\) rounds & \(p_0\) = 0.5 & \(p_2\) = 0.575 & \(P_O(i)\) \\
\midrule()
\endhead
6 & 3.1\% & 4.2\% & 4.7\% \\
7 & 9.4\% & 11.2\% & 11.9\% \\
8 & 16.4\% & 17.8\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.0\% & 21.5\% \\
11 & 24.6\% & 22.0\% & 23.7\% \\
\bottomrule()
\end{longtable}

@@

Despite the implication that one can model team round win probability in
CoD SnD as a constant, anecdotally many believe in momentum i.e.~streaks
not due to random chance. I now evaluate win frequency in rounds
following a streak of length \(k\). I account for the length of the
series (\(i\) rounds), modeling the expected round win probability after
a streak of length \(k\) as

\begin{equation}\protect\hypertarget{eq:pki}{}{
  \hat{P}(\text{win round | k-round winning streak, series lasting i rounds})=\{
  \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
  \end{array}.
}\label{eq:pki}\end{equation}

Miller et al. (2018) provide evidence for negative recency suggesting
that

\begin{equation}\protect\hypertarget{eq:pkid}{}{
  \begin{array}{rcl}
  \hat{P}_{ki,\text{diff}} & = & \hat{P}(\text{win round | k-round winning streak, series lasting i rounds}) -  \\
    & & \hat{P}(\text{lose round | k-round losing streak, series lasting i rounds}) \\
    E[\hat{P}_{ki,\text{diff}}] & \neq & 0
  \end{array}
}\label{eq:pkid}\end{equation}

given a deterministic i.i.d. Bernoulli probability
\(\hat{p}(\text{win round | k-round winning streak, series lasting i rounds)}\)
(Equation \ref{eq:pki}.\footnote{Note that these probabilites are
  separate from those in Section \textbackslash ref\{sec:analysis-1\}.}

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum (Carlson and Shu 2007), I focus on
round win percentages in rounds immediately following streaks of three
wins, \(k = 3\).\footnote{Three happens to also be a reasonable number
  for series that lasts at maximum 11 rounds.}

>>>>

The general formula for the probability \(P_E(r)\) that a best-of-\(s\)
series lasts \(r\) rounds given constant probability \(p\) of one
team\footnote{If \(p > 0.5\), then I might say that this team is the
  better team (known omnisciently).} winning each round is

@@

\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(r) = \frac{(r - 1)!}{s_1!(r - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq:series-length}\end{equation}

Table \ref{tbl:cod-prob-series-lasting-r-rounds} shows the expected
probabilities for \(s = 11\) under the naive assumption that
\(p_0 = 0.5\) for a given round, along with the observed proportions
\(P_O(r)\) in CoD SnD.

\begin{longtable}{rrrr}
\caption{The probabilities that a best-of-11 series lasts $r$ rounds ($P_E(r)$, where $r \in [6, 7, 8, 9, 10, 11]$) under the assumption that each team has a 50\% probability ($p_0 = 0.5$) of winning each game. The observed frequencies observed are shown as a count $N_O(r)$ and as a proportion $P_O(r)$ of all series.}\label{tbl:cod-prob-series-lasting-r-rounds} \\
\toprule
&  & \multicolumn{2}{c}{Observed} \\ 
\cmidrule(lr){3-4}
Series lasts $r$ rounds & $P_E(r)$ & $P_O(r)$ & $N_O(r)$ \\ 
\midrule
6 & $3.1\%$ & $4.7\%$ & 40 \\ 
7 & $9.4\%$ & $11.9\%$ & 101 \\ 
8 & $16.4\%$ & $16.5\%$ & 141 \\ 
9 & $21.9\%$ & $21.7\%$ & 185 \\ 
10 & $24.6\%$ & $21.5\%$ & 183 \\ 
11 & $24.6\%$ & $23.7\%$ & 202 \\ 
\bottomrule
\end{longtable}

@@

\begin{equation}\protect\hypertarget{eq:series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(r) - P_E(r))^2}{P_E(r)}, r \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}

@@

\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}

Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new
\(P_E(r)\) when re-applying Equation \ref{eq:series-length} for the
maximum likelihood estimate \(p_2 = 0.575\).\footnote{The method of
  moments and minimum \(\chi^2\) estimates for \(p\) are omitted simply
  because the results would be nearly identical to those for the maximum
  likelihood estimate of \(p\) (since they are all \(\approx 0.575\).} I
observe that \(P_E(r)\) is notably larger for \(p_{2}\) when
\(r \in [6, 7]\), more closely matching \(P_O(r)\). \(p_{2}\) is also
closer to \(P_O(r)\) for \(r \in [9, 10]\), although not for
\(r \in [8, 11]\).

\begin{longtable}[]{@{}rrrr@{}}
\caption{The probabilities that a best-of-11 series lasts $r$ rounds under the naive assumption $p_0 = 0.5$ and the maximum likelihood estimation $p_2 = 0.575$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
Series lasts \(r\) rounds & \(p_0(r)\) = 0.5 & \(p_2(r)\) = 0.575 & \(P_O(r)\) \\
\midrule()
\endhead
6 & 3.1\% & 4.2\% & 4.7\% \\
7 & 9.4\% & 11.2\% & 11.9\% \\
8 & 16.4\% & 17.8\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.0\% & 21.5\% \\
11 & 24.6\% & 22.0\% & 23.7\% \\
\bottomrule()
\end{longtable}

@@

Despite the implication that one can model team round win probability in
CoD SnD as a constant, anecdotally many believe in momentum i.e.~streaks
not due to random chance. I now evaluate win frequency in rounds
following a streak of length \(k\). I account for the length of the
series (\(r\) rounds), modeling the expected round win probability after
a streak of length \(k\) as

\begin{equation}\protect\hypertarget{eq:pwkr}{}{
  \hat{p}^{kr,\text{win}} = \hat{p}(\text{win round | k-round winning streak, series lasting r rounds})=\{
    \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
    \end{array}.
  }\label{eq:pwkr}
\end{equation}

\(\hat{p}^{kr,loss} = \hat{p}(\text{lose round | k-round losing streak, series lasting r rounds})\)
can be formulated symmetrically.

Miller et al. (2018) provide evidence for negative recency suggesting
that, given a deterministic i.i.d. Bernoulli probability (Equation
\ref{eq:pki},

\begin{equation}\protect\hypertarget{eq:pkrd}{}{
E[\hat{P}^{kr,\text{win}} - \hat{P}^{kr,\text{loss}}] = 0
}\label{eq:pkrd}\end{equation}

is not true due to streak-selection bias.\footnote{Note that these
  probabilities are separate from those in Section \ref{sec:analysis-1}.}
I don't attempt to measure this expectation like Miller et al. (2018),
who perform their analysis longitudinally per study participant;
nonetheless, I apply their formulation for applying the ``downward
bias'' on post-streak success to come up with estimates
\(\hat{P}_{kr,\text{win}}^{\text{adj.}}\), specifying \(r\) as the
number of trials.

Given that people typically perceive streaks as beginning after the
third success (or failure) at minimum (Carlson and Shu 2007), I focus on
round win percentages in rounds immediately following streaks of three
wins.\footnote{Three happens to also be a reasonable number for series
  that lasts at maximum 11 rounds.} Table
\ref{tbl:cod-round-win-prop-after-k-wins} shows
\(\hat{P}_{kr,\text{win}}\) for \(k = 3\) winning streaks, for both
teams that empirically won and lost the series. (Some combinations of
\(r\) and \(\text{team loses series}\) are irrelevant, such as the case
when \(r \in [7, 8, 9]\), so these combinations are excluded.)

\begin{longtable}{rrrrrrrr}
\toprule
&  &  & \multicolumn{3}{c}{$\hat{p}^{kr,\text{win}}$} & \multicolumn{2}{c}{$p$-value} \\
\cmidrule(lr){4-6} \cmidrule(lr){7-8}
Series lasts $r$ rounds & Win Series? & $N^{kr,\text{win}}_O$ & $O$ & $E$ & $\text{adj.}$ & $E$ & $\text{adj.}$ \\ 
\midrule
7 & win & 209 & $75\%$ & $75\%$ & $61\%$ & 0.58 & <=0.01 \\ 
8 & win & 209 & $62\%$ & $60\%$ & $45\%$ & 0.28 & <=0.01 \\ 
9 & win & 193 & $52\%$ & $50\%$ & $35\%$ & 0.33 & <=0.01 \\ 
10 & lose & 60 & $13\%$ & $14\%$ & $10\%$ & 0.64 & 0.26 \\ 
10 & win & 151 & $44\%$ & $43\%$ & $29\%$ & 0.45 & <=0.01 \\ 
11 & lose & 129 & $24\%$ & $25\%$ & $16\%$ & 0.63 & <=0.01 \\ 
11 & win & 150 & $40\%$ & $38\%$ & $25\%$ & 0.29 & <=0.01 \\ 
\bottomrule
\end{longtable}
